Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-27 review/review-response-01.tex

The two modes of falling are different. We have added more information in
section 2.4 to explain these two modes. If the participants were cycling on an
infinite plane and we perturb them to fall, then we would strictly use body
contact with the ground as the criteria to define a fall. We have now cited
research that indicates riders often attempt to remove their foot from the
pedal and place it on the ground to lessen or prevent a complete fall to the
ground, so using that as an indicator for defining a fall is justified. But we
have a limited width treadmill, so as speed increases a perturbation will cause
greater lateral deviations. If the wheel of the bicycle exits the width we had
to stop the treadmill for safety reasons. Neither of the fall modes on the
treadmill are ``extremely dangerous'' or ``dangerous'', based on our
experience, so we did not consider this aspect.

\begin{displayquote}
  According to the explanation of the experimental procedures, I understand
  that in this experiment, each force was tested only once, and the subjects
  were not allowed to fully familiarize themselves with the force. I believe
  that the interpretation of the results would be completely different if the
  number of trials with each force were small or large enough. The fact that
  the number of trials was analyzed as a significant variable in the regression
  also reflects this assumption. I would like you to add an explanation about
  this point in the experimental procedures or discussion section.
\end{displayquote}

The participants experienced the same magnitude perturbations more than once
throughout the warm-up, the perturbation threshold determination, and the
adaptive staircase procedure. But they did not experience the same force
repeatedly many times in order to gain familiarity. We would have liked to
remove the learning effect by letting the participants train for a long
duration, but this was not practical. We choose to include a measure of
learning in the statistics as an alternative. We have now noted this in Section
2.4.

@@

\begin{displayquote}
  The direction of the perturbation is said to be random from left to right,
  but I cannot decide whether it is appropriate to assume that it is truly
  random when analyzing the results. I believe that when the number of trials
  is small, the factors that affect the results are not only the magnitude of
  the force, but also the direction, which has a significant impact.
  Consideration and explanation is needed as to whether the results when a
  perturbation is in the opposite direction to the previous perturbation can be
  considered the same as the results when the perturbation is in the same
  direction. Furthermore, I presume that the results will be affected if the
  bicycle's posture at the time the perturbation is applied is leaning in the
  same direction as the perturbation, or if it is leaning in the opposite
  direction. If such conditions are taken into account in the analysis, please
  explain.
\end{displayquote}

Each participant experienced an equal number of clockwise and counterclockwise
perturbations. So 10 clockwise, 10 counterclockwise for both balance-assist on
and off, making 40 perturbations in total. Within the 20 perturbations of each
of the balance-assist states, the order of the directions was randomized.

In the analysis we normalized the time series measurements to the sign of the
perturbation as well as included the configuration at the time of perturbation
as independent variables. A variable could be included to indicate if a
perturbation was preceded by the same sign perturbation or not, but we have
chosen not to include this, as it was not in our original hypothesis. We added
text to clarify these things in Section 2.5.

### 2025-11-27 main.tex
Adjusted review response from Marten's comments.

\subsection{Protocol}
%
We recruited 26 able-bodied young adults between 20 and 36 years old (mean 26
years~\footnote{Excluding four participants' ages that were misplaced but that
did fall within the stated range.}) to participate in the experiments. The
participants were all confident in their cycling skills and had cycled at least
once in the last month. All participants consented to the experiment and could
decline to continue at any time. The study was approved by Delft University of
Technology's Human Research Ethics Council (record \#3897). The participants
were divided into two groups. The first group of fifteen participants
performed the protocol at a belt speed of 10~\si{\kph}(2.8~\si{\mps}) with the
gain factor set to \(\kappa=3.9\) and two weeks later the second group of
eleven participants performed the protocol at a belt speed of 6~\si{\kph}
(1.7~\si{\mps}) with the gain factor set to \(\kappa=5.2\).

@@

Following the initial threshold determination where the participant experienced
a range of perturbation magnitudes, we then choose perturbation forces
according to a random and adaptive staircase procedure applying perturbations
above and below the initial perturbation threshold, while allowing small
progression of the perturbation threshold to accommodate learning effects. We
did not try to eliminate learning effects due to the time per participant that
would be required, but factored this into the later statistical analysis.  Five
possible perturbation forces are determined based on the initial perturbation
threshold estimate: the initial estimate itself, two perturbations lower than
the initial estimate, and two perturbations higher than the initial estimate.
The five possible perturbations are separated by 10~\si{\newton} steps. For
example, if the initial estimate of the perturbation threshold of a participant
is 80~\si{\newton}, the five possible perturbations are 60, 70, 80, 90 and
100~\si{\newton}. Which one of these five perturbations is chosen first is
determined at random. If the perturbation results in a fall, the estimate of
the perturbation threshold is decreased by 10~\si{\newton}, and vice versa if
the perturbation did not result in a fall. Five new possible perturbation
forces are determined around the updated perturbation threshold, and a new
perturbation is selected at random. This process iterates until twenty
perturbations are applied. The goal of this adaptive staircase procedure is to
have participants fall for approximately 50\% of the time. After the first set
of 20 perturbations, we let the cyclist rest and then perform another 20
perturbations. We randomize whether the balance assist system is on or off
during the first or second set of 20 perturbations for each participant and the
participants were not told the order.

>>>>

\subsection{Protocol}
%
We recruited 26 able-bodied young adults between 20 and 36 years old (mean of
25 years) to participate in the experiments. The participants were all
confident in their cycling skills and had cycled at least once in the last
month. All participants consented to the experiment and could decline to
continue at any time. The study was approved by Delft University of
Technology's Human Research Ethics Council (record \#3897). The participants
were divided into two groups.  The first group of fifteen participants
performed the protocol at a belt speed of 10~\si{\kph}(2.8~\si{\mps}) with the
gain factor set to \(\kappa=3.9\) and two weeks later the second group of
eleven participants performed the protocol at a belt speed of 6~\si{\kph}
(1.7~\si{\mps}) with the gain factor set to \(\kappa=5.2\).

@@

Following the initial threshold determination where the participant experienced
a range of perturbation magnitudes, we then choose perturbation forces
according to a random and adaptive staircase procedure applying perturbations
above and below the initial perturbation threshold, while allowing small
progression of the perturbation threshold to accommodate learning effects. We
did not try to eliminate learning effects due to the time per participant that
would be required, but factored this into the later statistical analysis.  Five
possible perturbation forces are determined based on the initial perturbation
threshold estimate: the initial estimate itself, two perturbations lower than
the initial estimate, and two perturbations higher than the initial estimate.
The five possible perturbations are separated by 10~\si{\newton} steps. For
example, if the initial estimate of the perturbation threshold of a participant
is 80~\si{\newton}, the five possible perturbations are 60, 70, 80, 90 and
100~\si{\newton}. Which one of these five perturbations is chosen first is
determined at random. If the perturbation results in a fall, the estimate of
the perturbation threshold is decreased by 10~\si{\newton}, and vice versa if
the perturbation did not result in a fall. Five new possible perturbation
forces are determined around the updated perturbation threshold, and a new
perturbation is selected at random. This process iterates until twenty
perturbations are applied. The sense of the perturbation was randomized within
the 20 perturbations. The goal of this adaptive staircase procedure is to have
participants fall for approximately 50\% of the time. After the first set of
20 perturbations, we let the cyclist rest and then perform another 20
perturbations. We randomize whether the balance assist system is on or off
during the first or second set of 20 perturbations for each participant and the
participants were not told the order.
